---
title: "Alexander Albrecht"
author: "Data Science 2"
editor: visual
format:
  html:
    embed-resources: true
execute:
  echo: true
  message: false
  warning: false
---

```{r setup}
library(tidyverse)
library(tidytext)
library(schrute)
```

# The Office

Let's continue to work with *The Office* text from the [schrute](https://cran.r-project.org/web/packages/schrute/index.html) package.

# Sentiment Analysis

### Over Time

Calculate the sentiment score (positive - negative) for each episode, and plot this variable over time.

```{r}
library(tidyverse)
library(tidytext)
library(schrute)

# Data reminder
theoffice |> glimpse()

#Tokenize the text and join with sentiment lexicon
episode_sentiment <- theoffice |>
  mutate(episode_id = paste(season, episode, sep = "_")) |>
  unnest_tokens(word, text) |>
  inner_join(sentiments, by = "word") |>
  group_by(season, episode, episode_id) |>
  summarize(
    pos_score = sum(sentiment == "positive"),
    neg_score = sum(sentiment == "negative"),
    sent_score = pos_score - neg_score,
    sent_ratio = pos_score / neg_score,
    .groups = "drop"
  )

# Sequential order for episodes
episode_sentiment <- episode_sentiment |>
  arrange(season, episode) |>
  mutate(episode_num = row_number())

# Step 3: Plot sentiment over time
ggplot(episode_sentiment, aes(x = episode_num, y = sent_score)) +
  geom_line() +
  geom_point() +
  geom_smooth(method = "loess", se = FALSE, color = "blue") +
  labs(
    x = "Episode (chronological)",
    y = "Sentiment Score (Positive - Negative)",
    title = "Sentiment Over Time in The Office",
    subtitle = "Note: only 1 episode with a negative sentiment score: S3E20"
  )
```

### By Rating

Create a scatterplot of the `imdb_rating` versus the sentiment score (positive - negative), where the case is again an episode. Decide whether a linear model is a good fit to this relationship (optional, if you have had modeling classes, what model looks best?). Comment on what you find.

```{r}

# The theoffice data has imdb_rating - we need to join it with our sentiment
# First get unique episode info with ratings
episode_info <- theoffice |>
  distinct(season, episode, imdb_rating)

# Join with sentiment scores
sentiment_rating <- episode_sentiment |>
  left_join(episode_info, by = c("season", "episode"))

# Create scatterplot
ggplot(sentiment_rating, aes(x = sent_score, y = imdb_rating)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE, color = "blue") +
  labs(
    x = "Sentiment Score (Positive - Negative)",
    y = "IMDB Rating",
    title = "IMDB Rating vs. Sentiment Score",
    subtitle = "SE and Loess turned on but not necessarily statistically accurate, I just enjoy seeing them."
  )



# Check linear model fit
model <- lm(imdb_rating ~ sent_score, data = sentiment_rating)
summary(model)

# You could also check correlation
cor(sentiment_rating$sent_score, sentiment_rating$imdb_rating)

# 0.21 r correlation value, slight positive. 

```

Scatterplot looks positive but somewhat logistic in a sense. However a logistic model doesn't make statistical sense.

So, I'd say it's just a mild positive correlation that is very weak in the early data points.

**Andrea**: Great idea to run a linear model and check the correlation! This is something I didn't do in my assignment and it definitely adds weight to your argument. Since you mention that a logistic model doesn't make statistical sense, it would help to state what you visually observed, since this would make your statements stronger.

**Alexander**: Thank you! Explaining why logistical model doesn't make sense visually wouldl help. Great point. 

### Sentiment Negation

Identify at least a few negations that change the sentiment values based on the text. Use this to recreate the graphs above (or do this first, then create the graphs).

Claude helped with this problem heavily as I was confused by how to proceed.

**Andrea**: Great job being transparent about your process! To give the reviewer more context you should consider stating specifically what you were confused about and if there were any issues with Claude that you encountered along the way.

**Alexander**: Good reccomendation for future work. Worth being more clarifying about Claude. 

```{r}
# Define negation words
negation_words <- c("not", "no", "never", "without", "don't", "doesn't", 
                    "didn't", "won't", "wouldn't", "couldn't", "shouldn't",
                    "isn't", "aren't", "wasn't", "weren't", "can't")

# Step 1: Create bigrams (pairs of consecutive words)
office_bigrams <- theoffice |>
  mutate(episode_id = paste(season, episode, sep = "_")) |>
  unnest_tokens(bigram, text, token = "ngrams", n = 2) |>
  separate(bigram, c("word1", "word2"), sep = " ")

negated_sentiments <- office_bigrams |>
  filter(word1 %in% negation_words) |>
  # Join sentiment for the second word
  inner_join(sentiments, by = c("word2" = "word")) |>
  # Flip the sentiment
  mutate(sentiment = if_else(sentiment == "positive", "negative", "positive"))

# Step 3: Get regular (non-negated) sentiments
# First, identify which words are negated (to exclude them)
negated_words_in_context <- office_bigrams |>
  filter(word1 %in% negation_words) |>
  select(season, episode, index, word2) |>
  mutate(is_negated = TRUE)

negation_effect <- office_bigrams |>
  filter(word1 %in% negation_words) |>
  inner_join(sentiments, by = c("word2" = "word")) |>
  group_by(season, episode) |>
  summarize(
    # Original positives that are negated: should subtract 2 (remove +1, add -1)
    # Original negatives that are negated: should add 2 (remove -1, add +1)
    negation_adjustment = -2 * sum(sentiment == "positive") + 2 * sum(sentiment == "negative"),
    .groups = "drop"
  )

# Step 4: Apply adjustment to original sentiment scores
episode_sentiment_adjusted <- episode_sentiment |>
  left_join(negation_effect, by = c("season", "episode")) |>
  mutate(
    negation_adjustment = replace_na(negation_adjustment, 0),
    sent_score_adjusted = sent_score + negation_adjustment
  )

# RECREATE GRAPHS
ggplot(episode_sentiment_adjusted, aes(x = episode_num, y = sent_score_adjusted)) +
  geom_line() +
  geom_point() +
  geom_smooth(method = "loess", se = FALSE, color = "blue") +
  labs(
    x = "Episode (chronological)",
    y = "Sentiment Score (Positive - Negative)",
    title = "Sentiment Over Time in The Office",
    subtitle = "Note: only 1 episode with a negative sentiment score: S3E20"
  )


# Join with sentiment scores
sentiment_rating_adjusted <- episode_sentiment_adjusted |>
  left_join(episode_info, by = c("season", "episode"))

ggplot(sentiment_rating_adjusted, aes(x = sent_score_adjusted, y = imdb_rating)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE, color = "blue") +
  labs(
    x = "Sentiment Score (Positive - Negative)",
    y = "IMDB Rating",
    title = "IMDB Rating vs. Sentiment Score",
    subtitle = "SE and Loess turned on but not necessarily statistically accurate, I just enjoy seeing them."
  )
```

**Andrea**: One note for the "IMDB Rating vs. Sentiment Score" graph, it might be worth briefly explaining why SE and Loess aren't statistically appropriate here. That way it turns into more of an insight rather than just a personal preference.

**Alexander**: Great point. LOESS and SE aren't appropriate because they aren't true statistical models but more just help guide the eyes. 

```{r}
#SOME FUN ADJUSTMENTS:

# See which episodes had the biggest negation adjustments
episode_sentiment_adjusted |>
  mutate(abs_adjustment = abs(negation_adjustment)) |>
  arrange(desc(abs_adjustment)) |>
  select(season, episode, sent_score, negation_adjustment, sent_score_adjusted) |>
  head(15)

# Most positively adjsuted
episode_sentiment_adjusted |>
  arrange(desc(negation_adjustment)) |>
  select(season, episode, sent_score, negation_adjustment, sent_score_adjusted) |>
  head(15)

# Side-by-side comparison
episode_sentiment_adjusted |>
  select(episode_num, season, episode, sent_score, sent_score_adjusted) |>
  pivot_longer(cols = c(sent_score, sent_score_adjusted), 
               names_to = "type", 
               values_to = "score") |>
  ggplot(aes(x = episode_num, y = score, color = type)) +
  geom_line(alpha = 0.7) +
  labs(
    x = "Episode (chronological)",
    y = "Sentiment Score",
    title = "Original vs. Negation-Adjusted Sentiment",
    color = "Score Type"
  )

# How much did negation typically shift scores?
episode_sentiment_adjusted |>
  ggplot(aes(x = negation_adjustment)) +
  geom_histogram(binwidth = 5, fill = "steelblue", color = "white") +
  geom_vline(xintercept = 0, lty = 2) +
  labs(
    x = "Negation Adjustment",
    y = "Count",
    title = "Distribution of Negation Adjustments by Episode"
  )

# Summary stats
episode_sentiment_adjusted |>
  summarize(
    mean_adjustment = mean(negation_adjustment),
    median_adjustment = median(negation_adjustment),
    min_adjustment = min(negation_adjustment),
    max_adjustment = max(negation_adjustment)
  )

```

**Andrea**: The fun adjustments were very insightful! Another thing you could consider doing is identifying which main characters consistently drive positive or negative sentiment across episodes.

**Alexander**: Ooh, did not think about this. In a future project, this is worth doing. Thanks, Andrea!
